FictionEro2 - Determinants of Reality Beliefs

### 2025-05-29 study2/analysis/2_emotion.qmd
updated study1 rawdata
@DominiqueMakowski could you check whether the logic is alright, if so I will then re-run the analysis for study 1

test_bf(m1,m2, m3, m4, m5,)
compare_performance(m1, m2, m3, m4, m5, m6)

@@

```{r EMotionality Plot}
df$Emotionality <- as.numeric(normalize(predict(pca)$PC1))

@@

The results show that females consistently rate photographs as higher in emotionality compared to males. Additonally, males rate AI-Generated images higher in emotionality compared to photographs.

@@

Looking at **AI-attitudes, porn frequency and sexual activity**, and feedback as moderators, we can see that for females, emotionality ratings for AI-generated images are negatively moderated by positive attitudes towards AI, meaning that females with more positive AI attitudes rate the images as less emotional. For males, emotionality ratings are positively moderated by negative attitudes towards AI, meaning that males with more negative AI attitudes rate the images as more emotional.

Looking at **feedback** as a moderator, the results show that emotionality ratings for AI-generated images were significantly moderated only by positive attitudes toward AI in females and negative attitudes toward AI in males.

Interestingly, for females, higher positive attitudes (BAIT Positive) were associated with lower emotionality ratings of AI-generated images. For males, higher negative attitudes (BAIT Negative) were associated with higher emotionality ratings.

>>>>

test_bf(m1,m2, m3, m4, m5,m6)
compare_performance(m1, m2, m3, m4, m5, m6)

@@

```{r EMotionality Plot}
df$Emotionality <- as.numeric(normalize(-predict(pca)$PC1))

@@

The results show males rate AI-Generated images lower in emotionality compared to photographs.

@@

Looking at **AI-attitudes, porn frequency and sexual activity**, and feedback as moderators, we can see that for males reporting higher expertise in AI led to an increase of emotionality ratings for images labelled as AI-Generated. Additionally, males who reported having more negative attitiudes towards AI, rated AI-Generated images lower in emotionality compared to fotographs. While for feamles, those with more positive AI attitudes rate AI-generated images higher in emotionality.

Looking at **feedback** as a moderator, the results show that for males who reported finidng AI-generated images as more attractive rated these images higher in emotionality.Conversely, males who reported finding AI-generated images as less attractive rated these images lower in emotionality.

### 2025-05-29 study2/analysis/3_realness.qmd
updated study1 rawdata
@DominiqueMakowski could you check whether the logic is alright, if so I will then re-run the analysis for study 1

df <- df|>
  filter(SexualOrientation == "Heterosexual")
```
## Physical Characteristics

@@

>>>>

df <- df|>
  filter(SexualOrientation == "Heterosexual")
```

## Physical Characteristics

@@

### Association betwen Emotional responses and Realness Ratings

**Realness** ratings were positively correlated with **Arousal**, **Enticement**, and **Valence.**

To examine the unique role of **Arousal** in predicting **perceived realness**, a model was tested including both linear and quadratic effects of Arousal, along with its interactions with Gender and Condition (AI-generated vs. real images), modeling participants and images as random intercepts. 
To assess whether a non-linear (quadratic) component improved the model, a comparison was conducted between a model including a quadratic term for Arousal and one without. The model excluding the quadratic term was decisively favored, with Bayes Factors indicating that the data were over 1000 times more likely under the simpler linear model. This strongly supports the adequacy of a linear relationship between Arousal and Realness in this context. The following findings were seen in the simpler model:

- **Arousal** had a strong positive linear effect on Realness, indicating that higher arousal was **associated** with **increased perceptions of realness**.

- For **female** participants, the positive **effect of Arousal on Realness** is significantly **weaker** than for **males**. This suggests that women are less influenced by Arousal when making realness judgments.

- Among **males**, the effect of **Arousal, on realness ratings**, is significantly **weaker** when judging images previous labelled as **AI-generated** compared to real ones.

These results suggest a gender-specific modulation in how emotional arousal shapes beliefs about what is real, particularly when distinguishing between real and synthetic faces.

```{r}
df |>
  select(Gender, Arousal, Enticing, Valence, RealnessBelief, Realness) |>
  group_by(Gender) |>
  correlation::correlation()

p_a <- plot(cor_test(df, "Arousal", "Realness")) +
  ggplot2::facet_wrap(~Gender) +
  ggplot2::theme_bw()

p_e <- plot(cor_test(df, "Enticing", "Realness")) +
  ggplot2::facet_wrap(~Gender) +
  ggplot2::theme_bw()

p_v <- plot(cor_test(df, "Valence", "Realness")) +
  ggplot2::facet_wrap(~Gender) +
  ggplot2::theme_bw()

p_sub <- p_a/ p_e/ p_v

m1_dir <- make_model("Realness ~ Arousal / Gender/ Condition + (1 | Participant) + (1|Stimulus)", family=glmmTMB::ordbeta())
m2_dir <- make_model("Realness ~ poly(Arousal, 2) / Gender/ Condition + (1 | Participant) + (1|Stimulus)", family=glmmTMB::ordbeta())

compare_performance(m1_dir, m2_dir)
test_performance(m1_dir, m2_dir)
